---
title: "FDS 3.2 Data Wrangling Exercise 2"
author: "Oswaldo Salazar"
date: "March 20, 2016"
output: html_document
---

## Dealing with missing values

This exercise uses the data from the Titanic database.

### 0: Load the data in RStudio
Saved the data set as a CSV file called titanic_original.csv and loaded it in RStudio into a data frame.

```{r}
titanic_original <- read.csv("~/Documents/Personal Documents/Personal/R/3_2_FDS_Exercise/titanic_original.csv")
```

### 1: Port of embarkation

"The embarked column has one missing value, which is known to correspond to a passenger who actually embarked at Southampton. Find the missing value and replace it with S."

The original database has 1310 records. The last one has all NA values so I eliminated it using:

```{r}
titanic_data = titanic_original[-1310,]
```
 Then assgined the new database to a new one called titanic.
 
```{r}
titanic <- titanic_data
```
Then I realized the column embraked was a factor class, so I converted it into a character class.
```{r}
class(titanic$embarked)
titanic$embarked <- as.character(titanic$embarked)
```
After this, I changed all the blank values into "S".
```{r}
titanic$embarked <- ifelse(titanic$embarked == "", "S", titanic$embarked)
```

### 2: Age

2.1 Calculated the mean age value, excluding the NAs, then assign this value to the NAs.

```{r}
age_mean <- mean(titanic$age, na.rm = TRUE)
titanic$age <- ifelse(is.na(titanic$age), age_mean, titanic_data$age)
```

2.2 Other options would have been to choose a random value from 0 to 50, or to replace all with the same random value. Having the mean instead of any other value makes more sense since the missing values would probably also converge to this mean value.

### 3: Lifeboat

Changed all empty velues with NA.

```{r}
titanic$boat <- ifelse(titanic$boat == "", NA, titanic$boat)
```

### 4: Cabin

A large portion of the cabins was assigned to the members of the 1st class, whereas the 2nd or 3rd class passenger didn't have one, mostly.

Created a new column called has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r}
titanic$has_cabin_number <- ifelse(titanic$cabin != "", 1,0)
```

### 6: Submit the project on Github

```{r}
write.csv(titanic, file = "titanic_clean.csv")
```
